---
title: "Untitled"
output: html_document
date: "2026-02-23"
---

```{r}
library(dplyr)
```


```{r}
query <- 'American Generations'
wiki_urls <- textpress::fetch_wiki_urls(query)
wiki_urls
```


```{r}
# ugs <- "worldcat\\.org|x\\.com|doi\\.org|pubmed\\.ncbi\\.nlm\\.nih\\.gov"

# grepl("\\.(pdf|doc|docx|xls|xlsx|ppt|pptx|jpg|jpeg|png|gif|mp4|mp3|zip|txt|xml|doi)$",
      
      
full <- textpress::fetch_wiki_refs(wiki_urls[1:5]) |>
  data.table::rbindlist()
```


```{r}
web_urls <- textpress::fetch_urls(query, n_pages = 4)
```




```{r}
wiki_text <- wiki_urls[1:4] |> textpress::read_urls()
```



```{r}
wiki_refs_text <- ''
```



```{r}
web_text <- web_urls |>  
  filter(path_depth > 0) |> 
  pull(url) |> 
  textpress::read_urls()
```

```{r}
web_ss <- web_text |> 
  # mutate(doc_id = paste0(h1_title, '.', 
  #                        parent_heading, '.', node_id)) |>
  textpress::nlp_split_sentences(by = c('url', 'node_id'))
```


nlp_

```{r}
wiki_ss <- wiki_text |> 
  mutate(doc_id = paste0(h1_title, '.', 
                         parent_heading, '.', node_id)) |>
  textpress::nlp_split_sentences()
```


A familial generation is a group of living beings constituting a single step in the line of descent from an ancestor.[6] 
 
In developed nations the average familial generation length is in the high 20s and has even reached 30 years in some nations.[7] 

Factors such as greater industrialisation and demand for cheap labour, urbanisation, delayed first pregnancy and a greater uncertainty in both employment income and relationship stability have all contributed to the increase of the generation length from the late 18th century to the present.
