---
title: "Basic NLP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic NLP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

Three `nlp_*` functions form the core processing pipeline: `nlp_split_sentences()` segments text into sentence rows; `nlp_tokenize_text()` tokenizes each sentence into a named list with optional character spans (two methods: standard word-boundary and Biber-style); `nlp_cast_tokens()` flattens that list into a long-format data frame ready for counting, filtering, or indexing. This vignette walks through each step on a Wikipedia corpus, then collapses them into a single pipe.

```{r libs}
library(textpress)
library(dplyr)
library(DT)
```

## Corpus

Fetch Wikipedia articles on the Strauss-Howe generational theory and scrape the article bodies.

```{r corpus}
wiki_urls <- textpress::fetch_wiki_urls("Strauss Howe generational theory", limit = 5)

wiki_text_list <- textpress::read_urls(wiki_urls, exclude_wiki_refs = TRUE)

corpus <- wiki_text_list$text
nrow(corpus)
```



## Step 1 -- sentence split

`nlp_split_sentences()` segments each abstract into individual sentences. The output adds `sentence_id` and carries `start`/`end` character offsets for each sentence within the source text.

```{r split-sentences}
web_ss <- corpus |>
  textpress::nlp_split_sentences(by = c("doc_id", "node_id"))

web_ss |>
  slice(1:5) |>
  mutate(text = {
    words <- strsplit(text, "\\s+")
    sapply(words, function(w) paste(paste(w[seq_len(min(25, length(w)))], collapse = " "), "..."))
  }) |>
  DT::datatable(rownames = FALSE)
```

### Abbreviation handling

Sentence splitting fails at abbreviations -- a period after *"Sen."* or *"U.S."* looks like a sentence boundary to a naive splitter. `nlp_split_sentences()` handles this with a mask-then-restore approach: before splitting, dots in known abbreviations are swapped to underscores so stringi never sees them as boundaries; after splitting, they are restored. Single uppercase initials (`\b[A-Z]\.`) are protected the same way, covering constructions like *"George W. Bush"* without needing them in the list.

The built-in `abbreviations` vector covers titles (*Mr., Dr., Sen., Gov., Rep., Prof.*), months (*Jan. -- Dec.*), and common forms (*U.S., D.C., vs., i.e.*). It is short by design -- the regex for single-letter initials handles the long tail. The `abbreviations` parameter accepts any character vector, so domain-specific forms can be appended or the list replaced entirely.

```{r abbrevs}
textpress::abbreviations
```

```{r abbrevs-custom, eval=FALSE}
# Extend with domain-specific abbreviations
nlp_split_sentences(
  corpus,
  by            = c("doc_id", "node_id"),
  abbreviations = c(textpress::abbreviations, "approx.", "dept.", "est.")
)
```



## Step 2 -- tokenize

`nlp_tokenize_text()` tokenizes each sentence into a named list. With `include_spans = TRUE` the result carries two parallel lists -- `tokens` and `spans` -- keyed by a composite `uid` built from the `by` columns. Two tokenization methods are available via `method`: `"word"` (default) splits on word boundaries; `"biber"` applies a linguistically motivated scheme that splits contractions, handles hyphenated compounds, and follows Biber-style corpus tokenization conventions.

```{r tokenize}
tokens <- web_ss |>
  textpress::nlp_tokenize_text(
    by            = c("doc_id", "node_id", "sentence_id"),
    include_spans = TRUE
  )

# tokens list: first unit
head(tokens$tokens[[1]])

# spans matrix: matching start/end positions
head(tokens$spans[[1]])
```

### Biber method

The key difference is contraction handling. `"word"` treats *"they've"* as a single token; `"biber"` splits it into *"they"* + *"'ve"* -- and does the same for all common English contractions. Hyphenated compounds are also handled differently: hyphens within words are temporarily masked so *"well-known"* tokenizes as one unit rather than three.

```{r tokenize-biber}
tokens_biber <- web_ss |>
  textpress::nlp_tokenize_text(
    by            = c("doc_id", "node_id", "sentence_id"),
    include_spans = TRUE,
    method        = "biber"
  )

head(tokens_biber$tokens[[1]])
```



### tokenizers package

The `tokenizers` package provides a simple, fast baseline via `tokenize_words()`. It lowercases by default, strips punctuation, and returns a named list -- one character vector per input string. No spans, no contraction splitting, no biber logic, but useful for quick bag-of-words pipelines or as a reference point.

```{r tokenizers-eg}
eg <- c(
  "Sen. Sanders argued that they've long been ignored.",
  "The well-known theory spans four generational archetypes.",
  "U.S. voters aged 18-29 shifted dramatically from 2020 to 2024."
)

tokenizers::tokenize_words(eg)
```

The same sentences through the biber method -- note contraction splitting (*"they've"* → *"they"* + *"'ve"*), hyphen preservation (*"well-known"* as one token), and punctuation retention:

```{r tokenizers-vs-biber}
data.frame(text = eg, doc_id = seq_along(eg)) |>
  textpress::nlp_tokenize_text(
    by            = "doc_id",
    include_spans = FALSE,
    method        = "biber"
  )
```



## Step 3 -- cast to data frame

`nlp_cast_tokens()` flattens the token list into a long-format data table: one row per token, with `id`, `token`, `start`, and `end`.

```{r cast}
token_df <- tokens |> textpress::nlp_cast_tokens()

token_df |>
  slice(1:20) |>
  DT::datatable(rownames = FALSE)
```



## Full pipeline

All three steps as a single pipe from the abstract corpus.

```{r pipeline}
corpus |>
  textpress::nlp_split_sentences(by = c("doc_id", "node_id")) |>
  textpress::nlp_tokenize_text(
    by            = c("doc_id", "node_id", "sentence_id"),
    include_spans = TRUE
  ) |>
  textpress::nlp_cast_tokens()
```

## Summary

`nlp_split_sentences()` → `nlp_tokenize_text()` → `nlp_cast_tokens()` is the standard processing path from raw text to a token-level data frame. The `start`/`end` spans thread through each step, keeping token positions recoverable relative to the source text.
