---
title: "Basic NLP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic NLP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```

Three `nlp_*` functions form the core processing pipeline: `nlp_split_sentences()` segments text into sentence rows; `nlp_tokenize_text()` tokenizes each sentence into a named list with optional character spans; `nlp_cast_tokens()` flattens that list into a long-format data frame ready for counting, filtering, or indexing. This vignette walks through each step on a PubMed abstract corpus, then collapses them into a single pipe.

```{r libs}
library(textpress)
library(puremoe)
library(dplyr)
library(DT)
```

## Corpus

Search PubMed for abstracts on political ideology and fetch the records.

```{r corpus}
pmids <- puremoe::search_pubmed(
  '"political ideology"[TiAb]',
  retmax = 100
)

abstracts <- puremoe::get_records(
  pmids     = pmids,
  endpoint  = "pubmed_abstracts",
  cores     = 3
) |>
  filter(!is.na(abstract)) |>
  rename(doc_id = pmid, text = abstract) |>
  select(doc_id, year, journal, articletitle, text)

nrow(abstracts)
```



## Step 1 -- sentence split

`nlp_split_sentences()` segments each abstract into individual sentences. The output adds `sentence_id` and carries `start`/`end` character offsets for each sentence within the source text.

```{r split-sentences}
web_ss <- abstracts |>
  textpress::nlp_split_sentences(by = "doc_id")

web_ss |>
  slice(1:10) |>
  DT::datatable(rownames = FALSE)
```

### Abbreviation handling

Sentence splitting fails at abbreviations -- a period after *"Sen."* or *"U.S."* looks like a sentence boundary to a naive splitter. `nlp_split_sentences()` handles this with a mask-then-restore approach: before splitting, dots in known abbreviations are swapped to underscores so stringi never sees them as boundaries; after splitting, they are restored. Single uppercase initials (`\b[A-Z]\.`) are protected the same way, covering constructions like *"George W. Bush"* without needing them in the list.

The built-in `abbreviations` vector covers titles (*Mr., Dr., Sen., Gov., Rep., Prof.*), months (*Jan. -- Dec.*), and common forms (*U.S., D.C., vs., i.e.*). It is short by design -- the regex for single-letter initials handles the long tail. The `abbreviations` parameter accepts any character vector, so domain-specific forms can be appended or the list replaced entirely.

```{r abbrevs}
textpress::abbreviations
```

```{r abbrevs-custom, eval=FALSE}
# Extend with domain-specific abbreviations
nlp_split_sentences(
  corpus,
  by            = "doc_id",
  abbreviations = c(textpress::abbreviations, "approx.", "dept.", "est.")
)
```



## Step 2 -- tokenize

`nlp_tokenize_text()` tokenizes each sentence into a named list. With `include_spans = TRUE` the result carries two parallel lists -- `tokens` and `spans` -- keyed by a composite `uid` built from the `by` columns.

```{r tokenize}
tokens <- web_ss |>
  textpress::nlp_tokenize_text(
    by            = c("doc_id", "sentence_id"),
    include_spans = TRUE
  )

# tokens list: first unit
head(tokens$tokens[[1]])

# spans matrix: matching start/end positions
head(tokens$spans[[1]])
```



## Step 3 -- cast to data frame

`nlp_cast_tokens()` flattens the token list into a long-format data table: one row per token, with `id`, `token`, `start`, and `end`.

```{r cast}
token_df <- tokens |> textpress::nlp_cast_tokens()

token_df |>
  slice(1:20) |>
  DT::datatable(rownames = FALSE)
```



## Full pipeline

All three steps as a single pipe from the abstract corpus.

```{r pipeline}
abstracts |>
  textpress::nlp_split_sentences(by = "doc_id") |>
  textpress::nlp_tokenize_text(
    by            = c("doc_id", "sentence_id"),
    include_spans = TRUE
  ) |>
  textpress::nlp_cast_tokens()
```

## Summary

`nlp_split_sentences()` → `nlp_tokenize_text()` → `nlp_cast_tokens()` is the standard processing path from raw text to a token-level data frame. The `start`/`end` spans thread through each step, keeping token positions recoverable relative to the source text.
