---
title: "Regex search on a live corpus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Regex search on a live corpus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

With a corpus built and split into sentences, the simplest form of search is pattern matching. This vignette uses `search_regex()` across a generational-politics corpus to show what well-constructed regex actually buys you — age range constructions, turnout language, numeric change expressions — and how the results come back in a KWIC format ready for close reading or downstream aggregation.

## Search terms and web URLs

```{r search-terms}
search_terms <- c(
  "US Gen Z voters 2026",
  "US Millennial political party 2026",
  "US Gen X politics forgotten generation 2026",
  "US Baby Boomer Republican Democrat 2026"
)
```


```{r web-urls, message=FALSE, warning=FALSE}
library(textpress)
library(dplyr)
library(DT)

web_urls <- lapply(search_terms, function(x)
  textpress::fetch_urls(query = x, 
                        n_pages = 3, 
                        date_filter = 'm')) |>
  bind_rows() |>
  unique()
```



## Web text and sentence split

```{r web-text}
web_text_list <- web_urls |>
  filter(path_depth > 0) |>
  pull(url) |>
  textpress::read_urls(cores = 6) 

web_text <- web_text_list$text |>
  mutate(doc_id = match(url, unique(url))) |>
  relocate(doc_id, .before = 1)
```


```{r web-ss}
web_ss <- web_text |>
  textpress::nlp_split_sentences(by = c('doc_id', 'node_id'))
```



## Regex patterns

```{r patterns}
patterns <- list(
  energized = "\\b(?:energi[zs]ed|motivated|mobili[zs]ed|fired\\s+up|disillusioned|apathetic)\\b",
age_range = paste(
  "\\b(?:aged?(?:\\s+between)?|ages?)\\s+\\d{2,3}(?:(?:\\s*(?:[-–]|to|and)\\s*\\d{2,3})|(?:\\s+and\\s+(?:older|younger|over|under)))\\b",
  "\\bunder\\s+\\d{2}(?!\\d|,|%)\\b",
  "\\b\\d{2,3}(?!\\d|,|%)\\s*(?:\\+|and\\s+(?:older|younger|over|under))\\b",
  sep = "|"
),
  from_to   = "\\bfrom\\s+\\d+\\s+to\\s+\\d+\\b"
)
```



## Search results

### Search ~ `age_range`

```{r out-age-range}
fs <- web_ss |> 
  
  textpress::search_regex(
    query = patterns$age_range,
    by = c('doc_id', 'node_id'),
    highlight = c('<span style="background:#a6cbe1;">', '</span>')
    ) |>
  
  distinct(text, .keep_all = TRUE) |>
  select(doc_id, node_id, pattern, text)

if (!is.null(fs) && nrow(fs) > 0) fs |> DT::datatable(rownames = FALSE, escape = FALSE)
```



### Search ~ `from_to`

```{r out-from-to}
fs1 <- web_ss |> 
  
  textpress::search_regex(
    query = patterns$from_to,
    by = c('doc_id', 'node_id'),
    highlight = c('<span style="background:#fdc4a8;">', '</span>')
    ) |>
  
  distinct(text, .keep_all = TRUE) |>
  select(doc_id, node_id, pattern, text)

if (!is.null(fs1) && nrow(fs1) > 0) fs1 |> DT::datatable(rownames = FALSE, escape = FALSE)
```
