---
title: "Semantic search"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Semantic search}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
HUGGINGFACE_API_TOKEN <- Sys.getenv("HUGGINGFACE_API_TOKEN")
```


This vignette completes the search side of the textpress pipeline. Where regex and dictionary search match on form, and BM25 matches on term frequency, embedding-based search matches on meaning -- useful when the query and the relevant passage share little surface vocabulary. Here all three are run on the same generational-politics corpus, BM25 via `search_index()` and semantic via `search_vector()`, then combined with reciprocal rank fusion (RRF). The top-ranked chunks are passed to an LLM for synthesis and structured extraction, showing the full arc from raw URLs to generated output.

```{r message=FALSE, warning=FALSE}
library(textpress)
library(dplyr)
```



## Search terms and web URLs

```{r search-terms}
search_terms <- c(
  "US Gen Z voters 2026",
  "US Millennial political party 2026",
  "US Gen X politics 2026",
  "US Baby Boomer Republican Democrat 2026"
)
```



```{r web-urls}
web_urls <- lapply(search_terms, function(x)
  textpress::fetch_urls(query = x,
                        n_pages = 1,
                        date_filter = "m")) |>
  bind_rows() |>
  unique()
```

> 

## Web text and sentence split

`read_urls()` scrapes and parses the article text; `nlp_split_sentences()` segments each document into analysis-ready sentence rows.

```{r web-text}
web_text_list <- web_urls |>
  filter(path_depth > 0) |>
  pull(url) |>
  textpress::read_urls(cores = 4)

web_ss <- web_text_list$text |>
  textpress::nlp_split_sentences(by = c("doc_id", "node_id"))
```



## Chunks: 1 sentence + 1 sentence context

Roll each sentence into a chunk with one sentence of context (before/after) for retrieval. `nlp_roll_chunks()` with `chunk_size = 1` and `context_size = 1` yields a focal sentence plus surrounding context in `chunk_plus_context`.

```{r chunks}
chunks <- textpress::nlp_roll_chunks(
  web_ss,
  by = c("doc_id", "node_id", "sentence_id"),
  id_col = 'uid',
  chunk_size = 1,
  context_size = 1)
```



## Embeddings for chunks

Embed the chunk text (with context) via a Hugging Face inference endpoint. The result is a numeric matrix with one row per chunk; use it with `search_vector()` for semantic search.

```{r embeddings, eval=FALSE}
embeddings <- textpress::util_fetch_embeddings(
  chunks,
  by = "uid",
  api_token = HUGGINGFACE_API_TOKEN
)
dim(embeddings)
head(rownames(embeddings))
```



## BM25 index

Build a BM25 index over the same chunk text for keyword search. Tokenize by chunk, then index.

```{r index, eval=FALSE}
index <- textpress::nlp_tokenize_text(
  chunks,
  by = c("uid"),
  include_spans = FALSE
  ) |>
  textpress::nlp_index_tokens(stem = T)
```



## Search

A single query string is used for both BM25 and semantic search, enabling a direct comparison of the two methods on the same input.

```{r eval=FALSE}
query_text <- "Gen Z Millennial Gen X Boomer Republican Democrat party alignment 2026"
```


### BM25

Search the index with a short query relevant to the corpus (e.g. generations and voter turnout).

```{r search-bm25, eval=FALSE, message=FALSE, warning=FALSE}
bm25_hits <- search_index(index,
                          query_text,
                          n = 25,
                          stem = TRUE)

bm25_hits |>
  DT::datatable(
    rownames = FALSE,
    options = list(dom = "lrtip",
                   pageLength = 5))
```

### Semantic

Embed a query phrase that matches the corpus theme (generational politics, voter behavior), then run `search_vector()` and join back to chunks to show the retrieved text.

```{r search-semantic, eval=FALSE}
query_embed <- textpress::util_fetch_embeddings(
  query_text,
  api_token = HUGGINGFACE_API_TOKEN
)
semantic_hits <- search_vector(embeddings, query_embed, n = 25)

semantic_hits |>
  DT::datatable(
    rownames = FALSE,
    options = list(dom = "lrtip",
                   pageLength = 5))
```




### Hybrid: reciprocal rank fusion (RRF)

Merge the two ranked lists by *uid* and score each hit with `1 / (k + rank)`. *k* is a smoothing constant (here 60); larger *k* flattens the curve so lower ranks contribute more. Sum scores per *uid*, then sort descending. Chunks that appear in both lists get a higher fused score.

```{r rrf, eval=FALSE, message=FALSE, warning=FALSE}
k <- 60
both <- bind_rows(bm25_hits, semantic_hits) |>
  group_by(query, method) |>
  mutate(rank = row_number()) |>
  ungroup()

fused <- both |>
  mutate(rrf = 1 / (k + rank)) |>
  group_by(uid) |>
  summarise(rrf_score = round(100 * sum(rrf), 3), .groups = "drop") |>
  arrange(desc(rrf_score))

ff <- fused |>
  left_join(chunks) |>
  group_by(doc_id, node_id) |>
  slice_max(nchar(text), n = 1, with_ties = FALSE) |>
  ungroup() |>
  slice_max(rrf_score, n = 15)
```




## LLM calls over retrieved chunks

Format the top-ranked chunks as labeled source blocks and concatenate into a single context string for the LLM. Initialize the chat client.

```{r llm-setup, eval=FALSE}
library(ellmer)
meta <- web_text_list$meta |> select(doc_id, date)
ff_meta <- ff |> left_join(meta, by = "doc_id") |>
  mutate(
    header = if_else(!is.na(date) & nzchar(trimws(date)),
      paste0("## Source ", doc_id, " (", date, ")\n\n"),
      paste0("## Source ", doc_id, "\n\n")
    ),
    block = paste0(header, text)
  )
context <- paste(ff_meta$block, collapse = "\n\n---\n\n")
chat <- ellmer::chat_openai(model = "gpt-4o-mini")
```




### Generational portrait

Comparative structured portrait per generation (alignment, drivers, divisions).

```{r llm-portraits, eval=FALSE}
portrait_schema <- ellmer::type_object(
  "Comparative political portraits of each generation in 2026.",
  portraits = ellmer::type_array(
    ellmer::type_object(
      "Portrait of one generation.",
      generation               = ellmer::type_string("Gen Z / Millennial / Gen X / Boomer."),
      party_alignment          = ellmer::type_string("Current lean or alignment per the passages."),
      key_drivers              = ellmer::type_string("What is driving their alignment or behavior."),
      internal_divisions       = ellmer::type_string("Any splits within the generation mentioned."),
      contrast_with_prior      = ellmer::type_string("How they differ from the prior generation."),
      most_recent_source       = ellmer::type_string("Date of the most recent passage informing this portrait, as YYYY-MM-DD.")
    )
  )
)
portraits_out <- chat$chat_structured(
  glue::glue(
    "Write a comparative political portrait of each generation -- Gen Z,
   Millennials, Gen X, Boomers -- based solely on the passages.
   Focus on party alignment, key drivers, and internal divisions in 2026.

   Note the date of the most recent passage you draw on for each generation.

   PASSAGES:\n{context}"
  ),
  type = portrait_schema
)
portraits_out$portraits |> DT::datatable(rownames = F)
```

## Summary

`search_vector()` and `search_index()` cover complementary retrieval modes -- meaning vs. term frequency. Combining them with reciprocal rank fusion consistently outperforms either alone: chunks that rank well in both lists rise to the top regardless of surface vocabulary overlap. Passing the fused top chunks to an LLM closes the loop from raw URLs to structured output.







