---
title: "Semantic search"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Semantic search}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = "#>"
)
```


This vignette completes the search side of the textpress pipeline. Where regex and dictionary search match on form, and BM25 matches on term frequency, embedding-based search matches on meaning — useful when the query and the relevant passage share little surface vocabulary. Here all three are run on the same generational-politics corpus, BM25 via `search_index()` and semantic via `search_vector()`, then combined with reciprocal rank fusion (RRF). The top-ranked chunks are passed to an LLM for synthesis and structured extraction, showing the full arc from raw URLs to generated output.

```{r message=FALSE, warning=FALSE}
library(textpress)
library(dplyr)
library(ellmer)
```



## Search terms and web URLs

```{r search-terms}
search_terms <- c(
  "US Gen Z voters 2026",
  "US Millennial political party 2026",
  "US Gen X politics 2026",
  "US Baby Boomer Republican Democrat 2026"
)
```



```{r web-urls}
web_urls <- lapply(search_terms, function(x)
  textpress::fetch_urls(query = x,
                        n_pages = 1,
                        date_filter = "m")) |>
  bind_rows() |>
  unique()
```

> 

## Web text and sentence split

`read_urls()` scrapes and parses the article text; `nlp_split_sentences()` segments each document into analysis-ready sentence rows.

```{r web-text}
web_text_list <- web_urls |>
  filter(path_depth > 0) |>
  pull(url) |>
  textpress::read_urls(cores = 4)

web_ss <- web_text_list$text |>
  textpress::nlp_split_sentences(by = c("doc_id", "node_id"))
```



## Chunks: 1 sentence + 1 sentence context

Roll each sentence into a chunk with one sentence of context (before/after) for retrieval. `nlp_roll_chunks()` with `chunk_size = 1` and `context_size = 1` yields a focal sentence plus surrounding context in `chunk_plus_context`.

```{r chunks}
chunks <- textpress::nlp_roll_chunks(
  web_ss,
  by = c("doc_id", "node_id", "sentence_id"),
  id_col = 'uid',
  chunk_size = 1,
  context_size = 1)
```



## Embeddings for chunks

Embed the chunk text (with context) via a Hugging Face inference endpoint. The result is a numeric matrix with one row per chunk; use it with `search_vector()` for semantic search. Set `HUGGINGFACE_API_TOKEN` so the API call can run.

```{r}
embeddings <- textpress::util_fetch_embeddings(
  chunks,
  by = "uid",
  api_token = Sys.getenv("HUGGINGFACE_API_TOKEN")
)
dim(embeddings)
head(rownames(embeddings))
```



## BM25 index

Build a BM25 index over the same chunk text for keyword search. Tokenize by chunk, then index.

```{r index}
index <- textpress::nlp_tokenize_text(
  chunks,
  by = c("uid"),
  include_spans = FALSE
  ) |>
  textpress::nlp_index_tokens(stem = T)
```



## Search

```{r}
query_text <- "Gen Z Millennial Gen X Boomer Republican Democrat party alignment 2026"
```


### BM25

Search the index with a short query relevant to the corpus (e.g. generations and voter turnout).

```{r search-bm25, message=FALSE, warning=FALSE}
bm25_hits <- search_index(index,
                          query_text,
                          n = 25,
                          stem = TRUE)

bm25_hits |>
  DT::datatable(
    rownames = FALSE,
    options = list(dom = "lrtip",
                   pageLength = 5))
```

### Semantic

Embed a query phrase that matches the corpus theme (generational politics, voter behavior), then run `search_vector()` and join back to chunks to show the retrieved text.

```{r}
query_embed <- textpress::util_fetch_embeddings(
  query_text,
  api_token = Sys.getenv("HUGGINGFACE_API_TOKEN")
)

semantic_hits <- search_vector(embeddings, query_embed, n = 25)

semantic_hits |>
  DT::datatable(
    rownames = FALSE,
    options = list(dom = "lrtip",
                   pageLength = 5))
```




### Hybrid: reciprocal rank fusion (RRF)

Merge the two ranked lists by *uid* and score each hit with `1 / (k + rank)`. *k* is a smoothing constant (here 60); larger *k* flattens the curve so lower ranks contribute more. Sum scores per *uid*, then sort descending. Chunks that appear in both lists get a higher fused score.

```{r message=FALSE, warning=FALSE}
k <- 60
both <- bind_rows(bm25_hits, semantic_hits) |>
  group_by(query, method) |>
  mutate(rank = row_number()) |>
  ungroup()

fused <- both |>
  mutate(rrf = 1 / (k + rank)) |>
  group_by(uid) |>
  summarise(rrf_score = round(100 * sum(rrf), 3), .groups = "drop") |>
  arrange(desc(rrf_score))

ff <- fused |>
  left_join(chunks) |>
  group_by(doc_id, node_id) |>
  slice_max(nchar(text), n = 1, with_ties = FALSE) |>
  ungroup() |>
  slice_max(rrf_score, n = 15)
```




## LLM calls over retrieved chunks

Context and chat are set once; each example below uses them. Requires `ellmer` and `OPENAI_API_KEY`.

```{r llm-setup, eval=requireNamespace("ellmer", quietly=TRUE) && nzchar(Sys.getenv("OPENAI_API_KEY"))}
meta <- web_text_list$meta |> select(doc_id, date)
ff_meta <- ff |> left_join(meta, by = "doc_id") |>
  mutate(
    header = if_else(!is.na(date) & nzchar(trimws(date)),
      paste0("## Source ", doc_id, " (", date, ")\n\n"),
      paste0("## Source ", doc_id, "\n\n")
    ),
    block = paste0(header, text)
  )
context <- paste(ff_meta$block, collapse = "\n\n---\n\n")
chat <- ellmer::chat_openai(model = "gpt-4o-mini")
```





### Narrative synthesis (free-form)

Single open-ended prompt; model returns prose.

```{r llm-synthesis, eval=requireNamespace("ellmer", quietly=TRUE) && nzchar(Sys.getenv("OPENAI_API_KEY"))}
synthesis <- chat$chat(glue::glue(
  "You are a political analyst covering the 2026 US electoral landscape.
   Using ONLY the passages below, write a 3–4 paragraph synthesis that:
   1. Opens with the single most striking finding about how any generation —
      Gen Z, Millennials, Gen X, or Boomers — is aligning with or shifting
      away from the Republican or Democratic party heading into 2026.
   2. Explains what is driving those alignments — economic anxiety, cultural
      identity, candidate appeal, institutional trust, or other factors
      mentioned in the passages.
   3. Notes any tensions or surprises: where generational behavior defies
      expectations or where generations are splitting internally.
   4. Closes with what remains uncertain or contested about how generational
      politics will shape the 2026 elections.
   5. Where a finding is specific enough to warrant it, cite the source inline
   as (Source N, YYYY-MM-DD) immediately after the claim.
   
   Tone: sharp and analytical. Do NOT add information not in the passages.

   PASSAGES:\n{context}"
), echo = FALSE)
```

```{r}
cat('<div style="background:#f7f7f7; padding:1em; border-left:3px solid #ccc; font-size:0.95em;">',
    synthesis,
    '</div>')
```



### Causal mechanism map (structured)

Structured extraction of cause–effect pathways with labels and passage support.

```{r llm-mechanisms, eval=requireNamespace("ellmer", quietly=TRUE) && nzchar(Sys.getenv("OPENAI_API_KEY"))}
mechanism_schema <- ellmer::type_object(
  "Causal mechanisms driving generational party alignment in 2026.",
  mechanisms = ellmer::type_array(
    ellmer::type_object(
      "One causal pathway.",
      mechanism_name  = ellmer::type_string("Short label, e.g. 'Economic anxiety driving Millennial GOP shift'."),
      cause           = ellmer::type_string("The generational, cultural, or political factor."),
      effect          = ellmer::type_string("The party alignment or electoral behavior outcome."),
      mediators       = ellmer::type_string("Any intervening variables mentioned."),
      passage_support = ellmer::type_string("Brief quote or paraphrase from the passages."),
      source = ellmer::type_string("Source ID and date, e.g. 'Source 13, 2026-02-10'. Use 'Multiple' if drawn from more than one passage."
    )
  )
  )
)
  
mechanisms_out <- chat$chat_structured(
  glue::glue(
    "Identify every causal mechanism the passages invoke to explain why
     Gen Z, Millennials, Gen X, or Boomers are aligning with Republicans
     or Democrats in 2026. Consider economic, cultural, institutional,
     and candidate-driven explanations.
     
     For each mechanism, note the Source ID and date from the passage header.

     PASSAGES:\n{context}"
  ),
  type = mechanism_schema
)
mechanisms_out$mechanisms |> DT::datatable(rownames = F)
```




### Generational portrait (structured)

Comparative structured portrait per generation (alignment, drivers, divisions).

```{r llm-portraits, eval=requireNamespace("ellmer", quietly=TRUE) && nzchar(Sys.getenv("OPENAI_API_KEY"))}
portrait_schema <- ellmer::type_object(
  "Comparative political portraits of each generation in 2026.",
  portraits = ellmer::type_array(
    ellmer::type_object(
      "Portrait of one generation.",
      generation               = ellmer::type_string("Gen Z / Millennial / Gen X / Boomer."),
      party_alignment          = ellmer::type_string("Current lean or alignment per the passages."),
      key_drivers              = ellmer::type_string("What is driving their alignment or behavior."),
      internal_divisions       = ellmer::type_string("Any splits within the generation mentioned."),
      contrast_with_prior      = ellmer::type_string("How they differ from the prior generation."),
      most_recent_source = ellmer::type_string("Date of the most recent passage informing this portrait, as YYYY-MM-DD.")
    )
  )
)
portraits_out <- chat$chat_structured(
  glue::glue(
    "Write a comparative political portrait of each generation — Gen Z,
     Millennials, Gen X, Boomers — based solely on the passages.
     Focus on party alignment, key drivers, and internal divisions in 2026.
     
     Note the date of the most recent passage you draw on for each generation.

     PASSAGES:\n{context}"
  ),
  type = portrait_schema
)
portraits_out$portraits |> DT::datatable(rownames = F)
```







