---
title: "Web corpus to dictionary search: generations and co-occurrence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Web corpus to dictionary search: generations and co-occurrence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```



This vignette walks through a complete `textpress` pipeline — from raw web search to sentence-level dictionary matching to a PMI-weighted co-occurrence view — using recent news about US generational politics as the corpus. We **fetch** article URLs via `fetch_urls()`, **read and split** into sentences with `read_urls()` and `nlp_split_sentences()`, and **search** using built-in dictionaries (`dict_generations`, `dict_political`) via `search_dict()`.

```{r message=FALSE, warning=FALSE}
library(textpress)
library(dplyr)
library(DT)
```




## Search terms and web URLs

```{r search-terms}
search_terms <- c(
  "US Gen Z voters 2026",
  "US Millennial political party 2026",
  "US Gen X politics forgotten generation 2026",
  "US Baby Boomer Republican Democrat 2026"
)
```

```{r web-urls}
web_urls <- lapply(search_terms, function(x)
  textpress::fetch_urls(query = x, 
                        n_pages = 3, 
                        date_filter = "m")) |>
  bind_rows() |>
  unique()
```

## Web text and sentence split

`read_urls()` scrapes and parses the article text; `nlp_split_sentences()` segments each document into analysis-ready sentence rows.

```{r web-text}
web_text_list <- web_urls |>
  filter(path_depth > 0) |>
  pull(url) |>
  textpress::read_urls(cores = 4)

web_ss <- web_text_list$text |>
  mutate(doc_id = match(url, unique(url))) |>
  relocate(doc_id, .before = 1) |>
  textpress::nlp_split_sentences(by = c("doc_id", "node_id"))
```



## Dictionaries and one combined search

`textpress` ships with two curated dictionaries: `dict_generations` maps surface variants (e.g., *"Gen Z"*, *"Zoomers"*, *"iGen"*) to a canonical `TermName`; `dict_political` does the same for political identity labels (*"Democrat"*, *"progressive"*, *"MAGA"*, etc.). We show the political dictionary first, then stack both and run a single dictionary search.

```{r show-dict-political}
DT::datatable(
  textpress::dict_political,
  options = list(pageLength = 10),
  rownames = FALSE
)
```



Stack both dictionaries (with a category), pass all variants to `search_dict()` once, then join back to get standardized `TermName` and category. One search covers generation and political terms.

```{r stacked-dict}
dict_gen <- textpress::dict_generations |>
  mutate(variant_lc = tolower(variant), cat = "gen") |>
  select(variant, variant_lc, term_name = TermName, cat)
dict_pol <- textpress::dict_political |>
  mutate(variant_lc = tolower(variant), cat = "affil") |>
  select(variant, variant_lc, term_name = TermName, cat)
stacked_dict <- bind_rows(dict_gen, dict_pol)
```



```{r}
matches <- web_ss |>
  textpress::search_dict(
    by   = c("doc_id", "sentence_id"),
    terms = stacked_dict$variant
  ) |>
  left_join(
    stacked_dict |> select(variant_lc, term_name, cat),
    by = c("term" = "variant_lc")
  )
matches |> arrange(id) |> DT::datatable(rownames = F)
```



## Co-occurrence: counts and PMI (gen × affiliation)

Sentences that appear in both generation and affiliation matches form (gen, affil) pairs. We count pairs and compute PMI:

$$\text{PMI}(x, y) = \log \frac{N \cdot n_{xy}}{n_x \cdot n_y}$$

where $N$ is total pair count, $n_{xy}$ is joint count, and $n_x$, $n_y$ are marginal counts. 

```{r cooccur}
ggs   <- matches |> filter(cat == "gen")   |> distinct(id, gen = term_name)
affil <- matches |> filter(cat == "affil") |> distinct(id, affil = term_name)
ids_both <- intersect(ggs$id, affil$id)

pairs <- ggs |> filter(id %in% ids_both) |>
  inner_join(affil |> filter(id %in% ids_both), by = "id")

n_total  <- nrow(pairs)
count_xy <- pairs |> count(gen, affil, name = "n_xy")
count_x  <- pairs |> count(gen,   name = "n_gen")
count_y  <- pairs |> count(affil, name = "n_affil")

cooccur <- count_xy |>
  left_join(count_x, by = "gen") |>
  left_join(count_y, by = "affil") |>
  mutate(pmi = round(log(n_total * n_xy / (n_gen * n_affil)), 2)) |>
  select(gen, affil, n_xy, n_gen, n_affil, pmi) |>
  arrange(desc(n_xy))

DT::datatable(cooccur, options = list(pageLength = 15), rownames = FALSE)
```



## Visualize: PMI heatmap

Filtering to pairs with at least two co-occurrences reduces noise. Tile color encodes PMI; label is raw co-occurrence count. Not much doing.

```{r heatmap, fig.height=6, out.width="100%", fig.alt="PMI heatmap of generation x political affiliation co-occurrence"}
library(ggplot2)
plot_data <- cooccur |>
  filter(n_xy >= 2) |>
  mutate(
    gen   = reorder(gen,   n_gen),
    affil = reorder(affil, n_affil)
  )

ggplot(plot_data, aes(x = affil, y = gen)) +
  geom_tile(aes(fill = pmi), color = "white", linewidth = 0.6) +
  geom_text(aes(label = n_xy, size = n_xy), color = "white", fontface = "bold", show.legend = FALSE) +
  scale_fill_gradient2(
    low = "#ef8a62", mid = "#f7f7f7", high = "#67a9cf",
    midpoint = 0, name = "PMI"
  ) +
  scale_size_continuous(range = c(2.5, 5)) +
  labs(
    title = "Generations × political affiliation in recent US news",
    subtitle = "Tile color = PMI  |  Label = co-occurrence count  |  Pairs with n ≥ 2 only",
    x = NULL, y = NULL,
    caption = "Source: textpress web corpus · dict_generations + dict_political"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 35, hjust = 1),
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(color = "grey40"),
    legend.position = "right"
  )
```
