---
title: "Wikipedia to dictionary search: generations and co-occurrence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Wikipedia to dictionary search: generations and co-occurrence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette pulls Wikipedia articles and their reference URLs with `fetch_wiki_urls()` and `fetch_wiki_refs()`, reads article text with `read_urls()`, splits into sentences with `nlp_split_sentences()`, runs a dictionary search with `search_dict()` using `dict_generations`, then finds sentences where **Gen X** co-occurs with **political affiliation** language.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(textpress)
```

## Fetch Wikipedia URLs

```{r}
query <- 'American Generations'
wiki_urls <- textpress::fetch_wiki_urls(query)
wiki_urls
```

## Citation URLs (optional)

```{r}
# Optional: filter refs by domain or file type before reading, e.g.:
# grepl("worldcat\\.org|x\\.com|doi\\.org|pubmed\\.ncbi", url)
# grepl("\\.(pdf|doc|docx|xls|xlsx|ppt|pptx|jpg|jpeg|png|gif|mp4|mp3|zip|txt|xml|doi)$", url)

full <- textpress::fetch_wiki_refs(wiki_urls[1:5]) |>
  data.table::rbindlist()
```

## Read article text

```{r}
wiki_text <- wiki_urls[1:4] |> textpress::read_urls()
```

## Process: sentence split

```{r}
wiki_ss <- wiki_text |>
  mutate(doc_id = paste0(h1_title, '.', parent_heading, '.', node_id)) |>
  textpress::nlp_split_sentences(by = c('doc_id'))
```

## Dictionary search: generation terms

```{r}
ggs <- wiki_ss |> textpress::search_dict(
  by = c('doc_id', 'sentence_id'),
  terms = textpress::dict_generations$variant
)
ggs
```

## Dictionary search: political affiliation terms

```{r}
affil <- wiki_ss |> textpress::search_dict(
  by = c('doc_id', 'sentence_id'),
  terms = textpress::dict_political$variant
)
affil
```

## Normalize: join dicts back to get TermName

`search_dict` returns matched `term` (variant); join to the dicts to get standardized `TermName` for counts and PMI.

```{r}
dict_gen <- textpress::dict_generations |> mutate(variant_lc = tolower(variant))
dict_pol <- textpress::dict_political   |> mutate(variant_lc = tolower(variant))

ggs <- ggs |> left_join(select(dict_gen, variant_lc, gen = TermName), by = c("term" = "variant_lc"))
affil <- affil |> left_join(select(dict_pol, variant_lc, affil = TermName), by = c("term" = "variant_lc"))
```

## Stack both outputs

```{r}
stacked <- bind_rows(
  ggs  |> distinct(id, gen)  |> rename(term = gen)  |> mutate(cat = "gen"),
  affil |> distinct(id, affil) |> rename(term = affil) |> mutate(cat = "affil")
)
stacked |> arrange(id)
```

## Co-occurrence: counts and PMI (gen × affiliation)

Sentences that appear in both tables; build (gen, affil) pairs per sentence, then count and PMI.

```{r}
ids_both <- intersect(ggs$id, affil$id)

pairs <- ggs |> filter(id %in% ids_both) |> distinct(id, gen) |>
  inner_join(affil |> filter(id %in% ids_both) |> distinct(id, affil), by = "id")

n_total <- nrow(pairs)
count_xy <- pairs |> count(gen, affil, name = "n_xy")
count_x  <- pairs |> count(gen, name = "n_gen")
count_y  <- pairs |> count(affil, name = "n_affil")

cooccur <- count_xy |>
  left_join(count_x, by = "gen") |>
  left_join(count_y, by = "affil") |>
  mutate(pmi = log(n_total * n_xy / (n_gen * n_affil))) |>
  select(gen, affil, n_xy, n_gen, n_affil, pmi) |>
  arrange(desc(n_xy))

cooccur
```
